---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# zarrquet

Example netcdf created with

```R
gdalmdimtranslate -co FORMAT=NC4 -co ARRAY:COMPRESS=DEFLATE \
/vsicurl/https://thredds.nci.org.au/thredds/fileServer/gb6/BRAN/BRAN2020/daily/ocean_salt_1993_01.nc \
-scaleaxes "Time(4),st_ocean(4),yt_ocean(4),xt_ocean(4)" example4d.nc
```

```{r}
## ignore nv and st_edges_ocean
ncmeta::nc_dims("example4d.nc")

## as in python we have
#ds.salt.data.shape
#(7, 12, 375, 900)
#ds.salt.data.chunks
#(4, 6, 188, 450)

```


Virtualize in Python

```python
import virtualizarr
ds = virtualizarr.open_virtual_mfdataset("example4d.nc")
ds.attrs = {} ## bad encoding of attributes so we drop 
ds.virtualize.to_kerchunk("example4d.parquet", format = "parquet")
```

Also see how the chunks are keyed by chunk-in-array index: 

```
ds.salt.data.manifest.keys()
#dict_keys(['0.0.0.0', '0.0.0.1', '0.0.1.0', '0.0.1.1', '0.1.0.0', '0.1.0.1', '0.1.1.0', '0.1.1.1', '1.0.0.0', '1.0.0.1', '1.0.1.0', '1.0.1.1', '1.1.0.0', '1.1.0.1', '1.1.1.0', '1.1.1.1'])
```


Look at the virtual table (the default record size is 1e5 so I just isolate to the rows that are valid).  The 'path' is the literal file input in-situ (can be remapped with the API at creation time). 

https://virtualizarr.readthedocs.io/en/latest/generated/virtualizarr.accessor.VirtualiZarrDatasetAccessor.rename_paths.html#virtualizarr.accessor.VirtualiZarrDatasetAccessor.rename_paths

```{r}
d <- arrow::open_dataset("example4d.parquet/salt/refs.0.parq") |> dplyr::collect()
d[!is.na(d$path), ]

```

I strongly think the array key indexes should be in that table as well but hey, there's a strong move to Icechunk instead of this (or kerchunk json). 

The full encoding and metadata is in

```{r}
## we still get weird encoding from python->json so sub out inline NaN
str(jsonlite::fromJSON(gsub("NaN", "\"NaN\"", readr::read_file("example4d.parquet/.zmetadata"))))
```



